### 2024-04-02 MSc.tex

\documentclass[a4paper,12pt,twoside,BCOR=10mm]{scrbook}

% UoI MSc thesis template (English) V2.0.4 25.4.2022

% The license of this template is not fully clear, but: This template
% version is based on an earlier LaTeX template that was once available
% on a SENS UGLA page where the author and license are unknown.  As
% that version has obviously been put once online with the intend to be
% used by students, providing and using it should not be problem.

% Helmut Neukirchen https://uni.hi.is/helmut updated that earlier
% template using a tikzpicture-based approach for the title page
% created by Þór Arnar Curtis.

% The included logos are very likely copyrighted by the University of
% Iceland, but
% https://zeroheight.com/1b323cfd9/p/20f1e6-theses/b/29a61a
% states that the document is intended to make it easier for
% students, staff, and print-shops to format the thesis and ensure a
% standardised appearance. So providing them here and using them
% should be legal.

% Search below for "MODIFY THESE LINES ONLY": there you can enter your name, thesis title, etc.

% BibLaTeX is assumed for references. While Overleaf does this
% automatically for you, if you run it locally on the commandline,
% then you need to run first pdflatex MSc.tex, then biber MSc (without
% the .tex extension), and then again pdflatex MSc.tex

% Packages
\usepackage[utf8]{inputenc}
\usepackage[icelandic, english]{babel}
\usepackage{t1enc}
\usepackage{graphicx}
\usepackage[intoc]{nomencl}
\usepackage{enumerate,color}
\usepackage{url}
\usepackage[pdfborder={0 0 0}]{hyperref}
\BeforeTOCHead[toc]{\cleardoublepage\pdfbookmark{\contentsname}{toc}} % Add Table of Contents to PDF "bookmark" table of contents
\usepackage{appendix}
\usepackage{eso-pic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[sf,normalsize]{subfigure}
\usepackage[format=plain,labelformat=simple,labelsep=colon]{caption}
\usepackage{placeins}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{listings}

% Packages used for title page layout
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning}

\input{variables.tex}

% Blue color according to HÍ corporate design
\convertcolorspec{RGB}{16,9,159}{rgb}\tmphiblue
\definecolor{hiblue}{rgb}\tmphiblue

\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0cm}
\raggedbottom

\setkomafont{captionlabel}{\itshape}
\setkomafont{caption}{\itshape}
\setkomafont{section}{\FloatBarrier\Large}
\setcapwidth{\textwidth}
%\setcapwidth[l]{\textwidth} % The original template had the [l] which leads to a warning that it gets ignored, so to reduce warnings, removed it.
\setcapindent{1em}

\usepackage{lmodern} % Use Latin Modern (instead of the default Computer Modern that is rendered using a bitmap font).
\usepackage{fixcmex} % To fix that Latin Modern large symbol math fonts has by default only one size: https://tex.stackexchange.com/a/621536
% Times new roman font instead of the standard LaTeX fonts: has not been test -- try this on your own risk
%\usepackage[T1]{fontenc}
%\usepackage{mathptmx}

%%%%%%%%%%%%%%%%% Configurations (Useful defaults, but OK to change %%%%%%%%%%%%%%%%%%%
\graphicspath{{figs/}} % Figures in directory figs

% Bibliography

% \usepackage[sort&compress,authoryear]{natbib} % Uncoment if you want to used NatBib instead of BibLaTeX (and comment the bitlatex line below)

\usepackage{biblatex}  % BibLaTeX used for references. 
\usepackage{csquotes} % BibLaTex wants to have context sensitive quotes
\addbibresource{references.bib} %  Name of *.bib file containing references

%%%%%%%%%%% MODIFY THESE LINES ONLY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Some note on advisor(s) vs. thesis committee: At the School of
% Engineering and Natural Sciences, according to Regulation
% no. 994-2017
% https://english.hi.is/university_school_of_engineering_and_natural_sciences/regulation_no_994_2017_on_masters_study_at
% Article 5., a student has an administrative supervisor who is
% typically also the academic supervisor, i.e. supervising the thesis.
% If someone from outside is supervising, than the person becomes the
% academic supervisor and you have in addition the administrative
% supervisor from within HÍ.
% In addition, according to Article 7., there is a Master's degree
% committee that includes at least two persons, one of which shall be
% the student’s administrative supervisor.
% Despite these formally defined roles, it is somewhat a matter of
% taste whether you list all persons as supervisors or all as just
% thesis committee or list the one single academic supervisor as
% supervisor and then all persons (repeating the supervisor's name)
% as thesis committee.
% In the settings, that follow here, you can set this and many other settings.

\def\thesisyear{2024}       					% Year thesis submitted
\def\thesismonth{June}					% Month thesis submitted, e.g. "December"
\def\thesisauthor{Brynjar Geir Sigurðsson}				% Thesis authoreiningaraðferðinni
\def\thesistitle{A NN approach to predicting gust factors in complex landscape} % Title of thesis
%\def\thesissubtitle{XXSubtitle is rarely usedXX}		% Subtitle of thesis (optional)
\def\thesisshorttitle{} 	% Optional: if title of thesis is longer than 50 characters, it would not fit on the spine (kjölur) and in this case you, need to provide a short title here for the print shop. Otherwhise: make it empty
\def\thesiscredits{60 ECTS} 						% Credits awarded for the project
\def\thesissubject{Mechanical Engineering}
\def\thesiskind{M.Sc.}					% E.g. M.Sc. or Ph.D. thesis
\def\thesiskindformal{CCMagister Scientiarum}			% E.g. Magister Scientiarum
\def\thesisschool{School of Engineering and {Natural Sciences}}		% School
\def\thesisfaculty{Industrial Engineering, Mechanical Engineering and Computer Science}% Faculty name without "Faculty of" (gets added) 
\def\thesisaddress{Dunhagi 5}			        % Faculty or school office address street
\def\thesispostalcode{107}			                % Faculty or school office zip code
\def\thesistelephone{525 4000}					% Office telephone
%\def\thesispublisher{XX}					% Publisher (not used)
\def\thesissupervisors{Kristján Jónasson // Ólafur Pétur Pálsson}					% Names of supervisors (split by \\ if more than one).
\def\thesisnrofsupervisors{2}					% Number of supervisors (to use "Supervisor" vs. "Supervisors")
\def\thesiscommittee{Kristján Jónasson \\ Ólafur Pétur Pálsson}			% Thesis commitee must include the supervisor(s)
\def\thesisexaminer{XXNN3XX}				        % Examiner (prófdómari)
\def\thesisISBN{}           					% Thesis ISBN number (keep empty: not used anymore)
\def\thesisprinting{}						% Name of printsthop (keep empty if thesis is never printed)
%\def\thesisprinting{Háskólaprent, Fálkagata 2, 107 Reykjavík}	% Name of printsthop (keep empty if thesis is never printed)
\def\thesislicense{This thesis may not be copied in any form without author permission.} % Set license here (could also be some Creative Commons license)
%\def\thesiskeywords{Keyword1, Keyword2, Keyword3}		% Keywords (not used anywhere, hence commented out)

%%%%%%%%%%% STOP MODIFYING HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% Next modifications: search for "START MODIFYING HERE AGAIN" below %%%%%%%%%%

% We need this command if someone used \\ in the thesis title
\newcommand{\removelinebreaks}[1]{%
  \begingroup\def\\{}#1\endgroup}

\begin{document}
\hypersetup{pageanchor=false}
\pagenumbering{Alph} % To prevent page numer "1" to be used multiple times, used "A", "B", etc. for the first pages
\begin{titlepage} % This titlepage environment spans in fact multiple pages

% This is the cover title page. If you go to a print shop, they will
% ignore it and create their own cover page, i.e. this cover page here
% is only used by the PDF version that gets electronically archived.

  \thispagestyle{empty}

  % The banner at top and bottom (using a tikz overlay)
  \begin{tikzpicture}[remember picture,overlay]
    \node[anchor=north west, inner sep=0pt] at (current page.north west)
        {\includegraphics[width=\paperwidth]{banner}}; % The top banner (as a PNG) % TODO: A vector graphic would be better

    \node(bottom)[shape=rectangle, fill=hiblue, minimum height=10mm, minimum width=\paperwidth, anchor=south west] at (current page.south west) {}; % The bottom banner (a filled rectangle)

    \node[above=0.4cm of bottom] {
        \begin{tabular}{c} 
          \sffamily \small \textcolor{hiblue}{\textbf{\MakeUppercase{Faculty of \thesisfaculty{}}}}
        \end{tabular}
    };
  \end{tikzpicture}

  \enlargethispage{3cm}
  \vspace*{3.5cm} % Here starts the white space below the top banner

  % The centering used below is with respect to the page margins which
  % are not the same on left and right which prevents proper centering
  % with respect to the tikz centering (and the title page in
  % general).  Instead, we use a minipage that we shift horizontally
  % by 2.6 cm.  But minipage sets \parskip to 0, so we need to save
  % and restore it.  To be able to use vfill/stretch in a minipage,
  % the height needs to be specified: 20.0 cm.
  \newlength{\currentparskip}
  \setlength{\currentparskip}{\parskip}
  \hspace*{-2.6cm}
  \begin{minipage}[t][20.0cm]{1.0\paperwidth}
    \setlength{\parskip}{\currentparskip}
    \begin{center}
      \vspace*{ \stretch{1.5} }
      \huge \sffamily \bfseries \thesistitle{}

      %\normalfont \LARGE \sffamily \thesissubtitle{}

      \vspace{ \stretch{1.0} }
      \normalfont \Large \sffamily \thesisauthor{}

      \vspace*{ \stretch{2.75} }

      \thesismonth{}~\thesisyear{} % E.g. "March 2022"

      \vspace{ \stretch{0.5} }

      \normalfont \Large \sffamily {\thesiskind{}~thesis \\
      in \thesissubject{}}

      \vspace*{ \stretch{1.0} }
    \end{center}
  \end{minipage}

  \newpage

  \thispagestyle{empty} \mbox{} % This is the inside page of the cover (cover verso) which remains empty

  \newpage

  \thispagestyle{empty} % This is the inner title page
  \begin{center}
    \vspace*{ \stretch{0.5} }

    \Large \sffamily \bfseries \thesistitle{}

    %\normalfont \large \sffamily \thesissubtitle{}

    \vspace*{ \stretch{1.0} }

    \sffamily{\thesisauthor{}}

    \vspace*{ \stretch{1.0} }
    \normalsize \thesiscredits{}~ECTS thesis submitted in partial fulfillment of a \\
    \textit{\thesiskindformal{}} degree in \thesissubject{}
    \large

    \ifx\thesissupervisors\empty % Only print supervisor part if supervisor names are not empty
    \else
      \vspace*{ \stretch{1.0} }
      \ifnum\thesisnrofsupervisors>1 Supervisors \\
      \else Supervisor \\
      \fi
      \thesissupervisors{}
    \fi

    \ifx\thesiscommittee\empty % Only print thesis committee part if committee names are not empty
    \else
      \vspace*{ \stretch{0.25} }
      \thesiskind{}~Committee\\
      \thesiscommittee{}
    \fi

    \ifx\thesisexaminer\empty % Only print examiner part if thesisexaminer is not empty
    \else
      \vspace*{ \stretch{0.25} }
      Examiner \\
      \thesisexaminer
    \fi

    \vspace*{ \stretch{1.0} }

    Faculty of \thesisfaculty \\
    \thesisschool \\
    University of Iceland \\
    Reykjavik, \thesismonth~\thesisyear

    \vspace*{ \stretch{0.5} }
  \end{center}

  \newpage

  \thispagestyle{empty} % This is the title verso (colophon), i.e. imprint/copyright page
  \vspace*{\fill}
  % \setcounter{page}{0} \renewcommand{\baselinestretch}{1.5}\normalsize
  \sffamily{\removelinebreaks{\thesistitle}} \\
  \ifx\thesisshorttitle\empty % Show only if shorttitle is provided
  \else
  (\sffamily{\thesisshorttitle{}}) \\
  \fi
  %\sffamily{\removelinebreaks{\thesissubtitle{}}} \\

  \thesiscredits ~ECTS thesis submitted in partial fulfillment of a \thesiskind{}~degree in \thesissubject
\\ \\
  Faculty of \thesisfaculty \\
  \thesisschool \\
  University of Iceland \\
  \thesisaddress \\ 
  \thesispostalcode, Reykjavik 
  Iceland

  Telephone: \thesistelephone \\ \\ 
  \vspace*{\lineskip}

  Bibliographic information: \\
  \thesisauthor{} (\thesisyear{}) \emph{\removelinebreaks{\thesistitle{}}}, \thesiskind{}~thesis, Faculty of \thesisfaculty, University of Iceland.\\

  Copyright \textcopyright~\thesisyear~ \thesisauthor \\
  \thesislicense{}\\

  \ifx\thesisISBN\empty % Show only if ISBN is provided
  \else
  ISBN~\thesisISBN
  \fi

  \ifx\thesisprinting\empty % Show only if print shop is provided
  \else
  Printing: \thesisprinting \\
  \fi

  Reykjavik, Iceland, \thesismonth~\thesisyear \\

%%%%%%%%%%% START MODIFYING HERE AGAIN %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \newpage % Dedication page: remove completely if you have no dedication

  \thispagestyle{empty} \mbox{}

  \vfill

  \begin{center}
    \textit{
      To all the students who made the wise decision to use \LaTeX. % Replace this by your dedication
    }
  \end{center} \vspace*{5cm}

  \vfill

%%%%%%%%%%%%%%%%%%%%%%%%% If you have no dedication: remove until here %%%%%%%%%%%%%%%%%%%%

\end{titlepage}

\cleardoublepage

\pagenumbering{roman} % Abstract page

\setcounter{page}{5}

\setkomafont{section}{\huge} % The title "Abstract" and "Útfdráttur" should look like the chapters, i.e. use \huge (\chapter cannot be used as this would create a new page)
\section*{Abstract}
English abstract (ca. 250 words).

\vfill \vspace*{1cm}

\section*{Útdráttur}
Hér kemur útdráttur á íslensku sem er að hámarki 250 orð.
\vfill

\newpage % Empty page
\setkomafont{section}{\FloatBarrier\Large}
% The first "chapter" which will start on a new page
% Table of contents starts automatically on a right-hand side page ("recto").
% Table of contents, list of figures and tables are automatically generated by the commands below
\hypersetup{pageanchor=true}
\tableofcontents
\listoffigures
\listoftables

\chapter*{Abbreviations}
\addcontentsline{toc}{chapter}{Abbreviations}
\markboth{Abbreviations}{Abbreviations}

Í þessum kafla mega koma fram listar yfir skammstafanir og/eða breytuheiti. Gefið kaflanum nafn við hæfi, t.d.\ Skammstafanir eða Breytuheiti. Þessum kafla má sleppa ef hans er ekki þörf. \\

The section could be titled: Glossary, Variable Names, etc.

If you use acronyms, it is strongly recommended to use a LaTeX
acronyms package. For example, to use the package \emph{acronym}, add
in the preamble:
\begin{verbatim}
\usepackage{acronym}
\end{verbatim}
and then here in this chapter (to create the list of all acronyms), use:
\begin{verbatim}
\begin{acronym}[SENS]
  \acro{SENS}{School of Engineering and Natural Sciences}
  \acro{UoI}{University of Iceland}
\end{acronym}
\end{verbatim}
In the square bracket above, you need to put the longest acronym, so
that the list gets proper indentation based on the longest acronym.
Also note that you need to manually sort here that list.

Then, whereever you use the acronym in your text: \verb|\ac{UoI}|:
that will automatically expand the acronym at the first use, but use
only the short version after the first use.  (Use \verb|\acp| if need
the plural form. \verb|\acl| if you explicitly want to have the long
form only, \verb|\acf| if you explicitly want to have the full long
and short form, i.e.\ like first use of \verb|\ac|.)

Use \verb|\acresetall| to forget about earlier usage (=expansion) of
acronyms, e.g.\ if despite already used in Abstract or Introduction,
but you want to expand them later once again (starting from, e.g., in
Foundations chapter): add \verb|\acresetall| at start of Introduction
chapter (and maybe also again at start of Foundations chapter).

\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
Í þessum kafla koma fram þakkir til þeirra sem hafa styrkt rannsóknina með fjárframlögum, aðstöðu eða vinnu. t.d.\ styrktarsjóðir, fyrirtæki, leiðbeinendur, og aðrir aðilar sem hafa á einhvern hátt aðstoðað við gerð verkefnisins, þ.m.t.\ vinir og fjölskylda ef við á. Þakkir byrja á oddatölusíðu (hægri síðu).

\pagenumbering{arabic}
\setcounter{page}{1}

\include{Chapters/Chapter1}
\include{Chapters/Chapter2}
\include{Chapters/Chapter3}
\include{Chapters/Chapter4}

\end{document}