### 2025-01-29 Chapters/Chapter6.tex

\chapter{Discussion}
\label{Chapter6}
The goal of this tesis was to research whether it was possible to use reanalysis data to increase the predictability of wind gusts and see what influence including the landscape elevation would have. A quick statistical analysis will give a p-value of 0, that is these results are statistically significant. The improvement shown in Table (\ref{table:results}) is not chance. There is some pattern to be learned in the interval that leads to a higher predictability of wind gusts over baseline.