### 2025-05-09 MSc.tex
vinna

\documentclass[a4paper,12pt,twoside,BCOR=10mm,openany]{scrbook}

% UoI MSc thesis template (English) V2.0.4 25.4.2022

% The license of this template is not fully clear, but: This template
% version is based on an earlier LaTeX template that was once available
% on a SENS UGLA page where the author and license are unknown.  As
% that version has obviously been put once online with the intend to be
% used by students, providing and using it should not be problem.

% Helmut Neukirchen https://uni.hi.is/helmut updated that earlier
% template using a tikzpicture-based approach for the title page
% created by Þór Arnar Curtis.

% The included logos are very likely copyrighted by the University of
% Iceland, but
% https://zeroheight.com/1b323cfd9/p/20f1e6-theses/b/29a61a
% states that the document is intended to make it easier for
% students, staff, and print-shops to format the thesis and ensure a
% standardised appearance. So providing them here and using them
% should be legal.

% Search below for "MODIFY THESE LINES ONLY": there you can enter your name, thesis title, etc.

% BibLaTeX is assumed for references. While Overleaf does this
% automatically for you, if you run it locally on the commandline,
% then you need to run first pdflatex MSc.tex, then biber MSc (without
% the .tex extension), and then again pdflatex MSc.tex

% Packages
% BETRI TÖFLUR OG LISTAR
\usepackage{booktabs}         % Fyrir betri töflur
\usepackage{caption}          % M.a. aukabil á undan töflufyrirsögnum
\captionsetup{font=small,   % Smærra letur á mynda- og töflutexta,
              labelfont=bf, % "Mynd 1" og "Tafla 2" feitletrað 
              margin=2cm,
              skip = 0.5\baselineskip} 
\usepackage[utf8]{inputenc}
\usepackage[icelandic, english]{babel}
\usepackage{t1enc}
\usepackage{graphicx}
\usepackage[intoc]{nomencl}
\usepackage{enumerate,color}
\usepackage{url}
\usepackage[pdfborder={0 0 0}]{hyperref}
\BeforeTOCHead[toc]{\cleardoublepage\pdfbookmark{\contentsname}{toc}} % Add Table of Contents to PDF "bookmark" table of contents
\usepackage{appendix}
\usepackage{eso-pic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subcaption}
%\usepackage[sf,normalsize]{subfigure}
\usepackage[format=plain,labelformat=simple,labelsep=colon]{caption}
\usepackage{placeins}
%\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xspace}
\usepackage{float}
\usepackage{caption}
% Packages used for title page layout
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{arydshln}
\usepackage{acronym}
% Packages used for header vertical space along with the methods that set the spacing
% Adjust spacing before and after chapters
\RedeclareSectionCommand[
  beforeskip=-0pt, % Space before the chapter title
  afterskip=0pt    % Space after the chapter title
]{chapter}

% Adjust spacing before and after sections
\RedeclareSectionCommand[
  beforeskip=0pt,   % Space before the section title
  afterskip=1pt     % Space after the section title
]{section}

% Adjust spacing before and after subsections
\RedeclareSectionCommand[
  beforeskip=0pt,   % Space before the subsection title
  afterskip=1pt     % Space after the subsection title
]{subsection}

\usetikzlibrary{positioning}
\pgfkeys{/pgf/number format/.cd,fixed,precision=2}

% https://tex.stackexchange.com/questions/374661/keyword-highlighting-not-working-in-python-with-listings
\lstdefinestyle{Python}{
    language        = Python,
    basicstyle      = \ttfamily,
    keywordstyle    = \color{blue},
    keywordstyle    = [2] \color{teal}, % just to check that it works
    stringstyle     = \color{green},
    commentstyle    = \color{red}\ttfamily
}

\input{variables.tex}

% Blue color according to HÍ corporate design
\convertcolorspec{RGB}{16,9,159}{rgb}\tmphiblue
\definecolor{hiblue}{rgb}\tmphiblue

\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0cm}
\raggedbottom

\setkomafont{captionlabel}{\itshape}
\setkomafont{caption}{\itshape}
\setkomafont{section}{\FloatBarrier\Large}
\setcapwidth{\textwidth}
%\setcapwidth[l]{\textwidth} % The original template had the [l] which leads to a warning that it gets ignored, so to reduce warnings, removed it.
\setcapindent{1em}

\usepackage{lmodern} % Use Latin Modern (instead of the default Computer Modern that is rendered using a bitmap font).
\usepackage{fixcmex} % To fix that Latin Modern large symbol math fonts has by default only one size: https://tex.stackexchange.com/a/621536
% Times new roman font instead of the standard LaTeX fonts: has not been test -- try this on your own risk
%\usepackage[T1]{fontenc}
%\usepackage{mathptmx}

%%%%%%%%%%%%%%%%% Configurations (Useful defaults, but OK to change %%%%%%%%%%%%%%%%%%%
\graphicspath{{figs/}} % Figures in directory figs

% Bibliography

% \usepackage[sort&compress,authoryear]{natbib} % Uncoment if you want to used NatBib instead of BibLaTeX (and comment the bitlatex line below)

\usepackage{biblatex}  % BibLaTeX used for references. 
\usepackage{csquotes} % BibLaTex wants to have context sensitive quotes
\addbibresource{references.bib} %  Name of *.bib file containing references

%%%%%%%%%%% MODIFY THESE LINES ONLY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Some note on advisor(s) vs. thesis committee: At the School of
% Engineering and Natural Sciences, according to Regulation
% no. 994-2017
% https://english.hi.is/university_school_of_engineering_and_natural_sciences/regulation_no_994_2017_on_masters_study_at
% Article 5., a student has an administrative supervisor who is
% typically also the academic supervisor, i.e. supervising the thesis.
% If someone from outside is supervising, than the person becomes the
% academic supervisor and you have in addition the administrative
% supervisor from within HÍ.
% In addition, according to Article 7., there is a Master's degree
% committee that includes at least two persons, one of which shall be
% the student’s administrative supervisor.
% Despite these formally defined roles, it is somewhat a matter of
% taste whether you list all persons as supervisors or all as just
% thesis committee or list the one single academic supervisor as
% supervisor and then all persons (repeating the supervisor's name)
% as thesis committee.
% In the settings, that follow here, you can set this and many other settings.

\def\thesisyear{2025}       					% Year thesis submitted
\def\thesismonth{June}					% Month thesis submitted, e.g. "December"
\def\thesisauthor{Brynjar Geir Sigurðsson}				% Thesis authoreiningaraðferðinni
\def\thesistitle{Prediction of Gust Factor in Complex Landscape with Neural Networks} % Title of thesis
%\def\thesissubtitle{XXSubtitle is rarely usedXX}		% Subtitle of thesis (optional)
\def\thesisshorttitle{} 	% Optional: if title of thesis is longer than 50 characters, it would not fit on the spine (kjölur) and in this case you, need to provide a short title here for the print shop. Otherwhise: make it empty
\def\thesiscredits{60} 						% Credits awarded for the project
\def\thesissubject{Mechanical Engineering}
\def\thesiskind{M.Sc.}					% E.g. M.Sc. or Ph.D. thesis
\def\thesiskindformal{Magister Scientiarum}			% E.g. Magister Scientiarum
\def\thesisschool{School of Engineering and {Natural Sciences}}		% School
\def\thesisfaculty{Industrial Engineering, Mechanical Engineering and Computer Science}% Faculty name without "Faculty of" (gets added) 
\def\thesisaddress{Dunhagi 5}			        % Faculty or school office address street
\def\thesispostalcode{107}			                % Faculty or school office zip code
\def\thesistelephone{525 4000}					% Office telephone
%\def\thesispublisher{XX}					% Publisher (not used)
\def\thesissupervisors{Kristján Jónasson}					% Names of supervisors (split by \\ if more than one).
\def\thesisnrofsupervisors{1}					% Number of supervisors (to use "Supervisor" vs. "Supervisors")
\def\thesiscommittee{Kristján Jónasson \\ Ólafur Pétur Pálsson \\ Guðrún Nína Petersen}			% Thesis commitee must include the supervisor(s)
\def\thesisexaminer{Halldór Björnsson}				        % Examiner (prófdómari)
\def\thesisISBN{}           					% Thesis ISBN number (keep empty: not used anymore)
\def\thesisprinting{}						% Name of printsthop (keep empty if thesis is never printed)
%\def\thesisprinting{Háskólaprent, Fálkagata 2, 107 Reykjavík}	% Name of printsthop (keep empty if thesis is never printed)
\def\thesislicense{This thesis may not be copied in any form without author permission.} % Set license here (could also be some Creative Commons license)
%\def\thesiskeywords{Keyword1, Keyword2, Keyword3}		% Keywords (not used anywhere, hence commented out)

%%%%%%%%%%% STOP MODIFYING HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% Next modifications: search for "START MODIFYING HERE AGAIN" below %%%%%%%%%%

% We need this command if someone used \\ in the thesis title
\newcommand{\removelinebreaks}[1]{%
  \begingroup\def\\{}#1\endgroup}

% To be able to use units in equations. From https://qerub.se/typesetting-units-in-latex
\newcommand{\unit}[1]{\ensuremath{\, \mathrm{#1}}}

\begin{document}
\hypersetup{pageanchor=false}
\pagenumbering{Alph} % To prevent page numer "1" to be used multiple times, used "A", "B", etc. for the first pages
\begin{titlepage} % This titlepage environment spans in fact multiple pages

% This must be placed after "\begin{document}":
\lstset{
    frame       = single,
    numbers     = left,
    showspaces  = false,
    showstringspaces    = false,
    captionpos  = t,
    caption     = \lstname
}

% This is the cover title page. If you go to a print shop, they will
% ignore it and create their own cover page, i.e. this cover page here
% is only used by the PDF version that gets electronically archived.

  \thispagestyle{empty}

  % The banner at top and bottom (using a tikz overlay)
  \begin{tikzpicture}[remember picture,overlay]
    \node[anchor=north west, inner sep=0pt] at (current page.north west)
        {\includegraphics[width=\paperwidth]{banner}}; % The top banner (as a PNG) % TODO: A vector graphic would be better

    \node(bottom)[shape=rectangle, fill=hiblue, minimum height=10mm, minimum width=\paperwidth, anchor=south west] at (current page.south west) {}; % The bottom banner (a filled rectangle)

    \node[above=0.4cm of bottom] {
        \begin{tabular}{c} 
          \sffamily \small \textcolor{hiblue}{\textbf{\MakeUppercase{Faculty of \thesisfaculty{}}}}
        \end{tabular}
    };
  \end{tikzpicture}

  \enlargethispage{3cm}
  \vspace*{3.5cm} % Here starts the white space below the top banner

  % The centering used below is with respect to the page margins which
  % are not the same on left and right which prevents proper centering
  % with respect to the tikz centering (and the title page in
  % general).  Instead, we use a minipage that we shift horizontally
  % by 2.6 cm.  But minipage sets \parskip to 0, so we need to save
  % and restore it.  To be able to use vfill/stretch in a minipage,
  % the height needs to be specified: 20.0 cm.
  \newlength{\currentparskip}
  \setlength{\currentparskip}{\parskip}
  \hspace*{-2.6cm}
  \begin{minipage}[t][20.0cm]{1.0\paperwidth}
    \setlength{\parskip}{\currentparskip}
    \begin{center}
      \vspace*{ \stretch{1.5} }
      \huge \sffamily \bfseries \thesistitle{}

      %\normalfont \LARGE \sffamily \thesissubtitle{}

      \vspace{ \stretch{1.0} }
      \normalfont \Large \sffamily \thesisauthor{}

      \vspace*{ \stretch{2.75} }

      \thesismonth{}~\thesisyear{} % E.g. "March 2022"

      \vspace{ \stretch{0.5} }

      \normalfont \Large \sffamily {\thesiskind{}~thesis \\
      in \thesissubject{}}

      \vspace*{ \stretch{1.0} }
    \end{center}
  \end{minipage}

  \newpage

  \thispagestyle{empty} \mbox{} % This is the inside page of the cover (cover verso) which remains empty

  \newpage

  \thispagestyle{empty} % This is the inner title page
  \begin{center}
    \vspace*{ \stretch{0.5} }

    \Large \sffamily \bfseries \thesistitle{}

    %\normalfont \large \sffamily \thesissubtitle{}

    \vspace*{ \stretch{1.0} }

    \sffamily{\thesisauthor{}}

    \vspace*{ \stretch{1.0} }
    \normalsize \thesiscredits{}~ECTS thesis submitted in partial fulfillment of a \\
    \textit{\thesiskindformal{}} degree in \thesissubject{}
    \large

    \ifx\thesissupervisors\empty % Only print supervisor part if supervisor names are not empty
    \else
      \vspace*{ \stretch{1.0} }
      \ifnum\thesisnrofsupervisors>1 Supervisors \\
      \else Supervisor \\
      \fi
      \thesissupervisors{}
    \fi

    \ifx\thesiscommittee\empty % Only print thesis committee part if committee names are not empty
    \else
      \vspace*{ \stretch{0.25} }
      \thesiskind{}~Committee\\
      \thesiscommittee{}
    \fi

    \ifx\thesisexaminer\empty % Only print examiner part if thesisexaminer is not empty
    \else
      \vspace*{ \stretch{0.25} }
      Examiner \\
      \thesisexaminer
    \fi

    \vspace*{ \stretch{1.0} }

    Faculty of \thesisfaculty \\
    \thesisschool \\
    University of Iceland \\
    Reykjavik, \thesismonth~\thesisyear

    \vspace*{ \stretch{0.5} }
  \end{center}

  \newpage

  \thispagestyle{empty} % This is the title verso (colophon), i.e. imprint/copyright page
  \vspace*{\fill}
  % \setcounter{page}{0} \renewcommand{\baselinestretch}{1.5}\normalsize
  \sffamily{\removelinebreaks{\thesistitle}} \\
  \ifx\thesisshorttitle\empty % Show only if shorttitle is provided
  \else
  (\sffamily{\thesisshorttitle{}}) \\
  \fi
  %\sffamily{\removelinebreaks{\thesissubtitle{}}} \\

  \thesiscredits ~ECTS thesis submitted in partial fulfillment of a \thesiskind{}~degree in \thesissubject
\\ \\
  Faculty of \thesisfaculty \\
  \thesisschool \\
  University of Iceland \\
  \thesisaddress \\ 
  \thesispostalcode, Reykjavik 
  Iceland

  Telephone: \thesistelephone \\ \\ 
  \vspace*{\lineskip}

  Bibliographic information: \\
  \thesisauthor{} (\thesisyear{}) \emph{\removelinebreaks{\thesistitle{}}}, \thesiskind{}~thesis, Faculty of \thesisfaculty, University of Iceland.\\

  Copyright \textcopyright~\thesisyear~ \thesisauthor \\
  \thesislicense{}\\

  \ifx\thesisISBN\empty % Show only if ISBN is provided
  \else
  ISBN~\thesisISBN
  \fi

  \ifx\thesisprinting\empty % Show only if print shop is provided
  \else
  Printing: \thesisprinting \\
  \fi

  Reykjavik, Iceland, \thesismonth~\thesisyear \\

\end{titlepage}

\cleardoublepage

\pagenumbering{roman} % Abstract page

\setcounter{page}{5}

\setkomafont{section}{\huge} % The title "Abstract" and "Útdráttur" should look like the chapters, i.e. use \huge (\chapter cannot be used as this would create a new page)

\section*{Abstract}
Whether it is wind farms, traffic accidents, high-voltage towers, or other things affected by wind, accurately predicting wind loads is crucial for reducing damage, improving operational safety, and boosting productivity. Wind gusts are multi-second spikes in speed, and wind gust factor is defined as the ratio of the highest three-second gust to the ten-minute mean wind speed. It is important to be able to predict both wind and gust factor accurately. Historically, numerical weather prediction models based on physical equations have been used to predict such variables. In recent years, artificial intelligence has advanced rapidly, and AI-based models have begun to rival conventional model performance. This project combines both approaches by using atmospheric model reanalysis output and digital elevation data as inputs to a deep neural network. Selected features from the reanalysis fields, terrain elevation, and derived variables are fed into several network architectures, whose performance is then compared against a simple linear regression baseline. The results are very promising: this hybrid methodology can substantially improve both instantaneous estimation and short-term forecasting of gust factors, particularly in areas of complex terrain.
\vfill \vspace*{1cm}

\section*{Útdráttur}
Hvort sem um er að ræða vindmyllugarða, umferðarslys, háspennumöstur, eða aðra hluti sem vindur getur haft áhrif á er spágeta um álag vegna vindsins mikilvæg, til að minnka eða koma í veg fyrir tjón, auka rekstraröryggi, eða bæta framleiðni. Vindhviður eru nokkurra sekúndna aukning á vindi og hviðustuðull mælir hæsta hlutfall milli þriggja sekúndna hviðu og 10 mínútna meðalvinds. Góðar hviðuspár eru mikilvægar. Sögulega hefur verið notast við hefðbundin spálíkön sem reiða sig á töluleg eðlisfræðileg líkön til að spá fyrir um veður, þar með talið hviðustuðul. Á síðustu árum hefur gervigreind þróast mikið og líkön verið þjálfuð til þess að spá fyrir um veður. Svo miklar framfarir hafa orðið að sum líkön geta keppt við hefðbundin veðurlíkön. Markmið þessa verkefnis er reyna að nýta bæði hefðbundin veðurlíkön og gervigreind til þess að meta hviðustuðla. Þetta er gert með því að nota endurgreiningargögn úr hefðbundnu veðurlíkani sem grunngögn í gervigreindarlíkan. Einnig eru notað hæðarlíkan sem lýsir landslagi. Djúpt tauganet er búið til með því að velja breytur úr endurgreiningargögnunum og hæðarlíkaninu og afleiddum breytum. Nokkur líkön af þessu tagi eru svo þjálfuð og niðurstöður bornar saman við einfalt grunnlíkan sem fengið er með línulega aðhvarfi. Niðurstöðurnar lofa mjög góðu um að þessi aðferðafræði geti bætt bæði samtímamat á hviðustuðli og spár um hann umtalsvert. Þetta á ekki síst við í flóknu landslagi.
\vfill

\newpage % Empty page
\setkomafont{section}{\FloatBarrier\Large}
% The first "chapter" which will start on a new page
% Table of contents starts automatically on a right-hand side page ("recto").
% Table of contents, list of figures and tables are automatically generated by the commands below
\hypersetup{pageanchor=true}
\tableofcontents
\listoffigures
\listoftables
\lstlistoflistings

\chapter*{Abbreviations}
\addcontentsline{toc}{chapter}{Abbreviations}
\markboth{Abbreviations}{Abbreviations}
\begin{acronym}[GeoTIFF]
  \acro{API}{Application Programming Interface}
  \acro{ASL}{Above Sea Level}
  \acro{AWS}{Automatic Weather Stations}
  \acro{CARRA}{Copernicus Artic Regional ReAnalysis dataset}
  \acro{CNN}{Convolutional Neural Networks}
  \acro{DEM}{Digital Elevaiton Model}
  \acro{ELI5}{Explain Like I am 5}
  \acro{ECMWF}{European Centre for Medium-Range Weather Forecast}
  \acro{GCM}{General Circulation Model}
  \acro{GeoTIFF}{Georeferenced TIFF}
  \acro{GPU}{Graphical Processing Unit}
  \acro{HRES}{High Resolution forecas}
  \acro{IMO}{Icelandic Meteoroligcal Office}
  \acro{IRCA}{Icelandic Road and Coastal Administration}
  \acro{JNWPU}{Joint Numerical Weather Prediction Unit}
  \acro{LWM}{Large AI Weather forecast Model}
  \acro{MAE}{Mean Asbolute Error}
  \acro{MAPE}{Mean Absolute Percentage Error}
  \acro{NN}{Neural Network}
  \acro{NWP}{Numerical Weather Prediction}
  \acro{SENS}{School of Engineering and Natural Sciences}
  \acro{TIFF}{Tag Image File Format}
  \acro{UoI}{University of Iceland}
\end{acronym}

\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
Special thanks go to my advisor, Kristján Jónasson, who helped me with the thesis work and sat with me to go over the code. Thanks also to Ólafur Pétur Pálsson who provided detailed notes and offered very helpful suggestions. I am also grateful to Guðrún Nína Petersen, who provided valuable insight into meteorological matters and gave constructive feedback on the thesis.

Finally, I would like to thank my parents without whom I would probably never had any want to finish.

\pagenumbering{arabic}
\setcounter{page}{1}

\include{Chapters/Chapter1}
\include{Chapters/Chapter2}
\include{Chapters/Chapter3}
\include{Chapters/Chapter4}

\printbibliography
\appendix
\include{Appendicies/AppendixA}
\include{Appendicies/AppendixB}

\end{document}