### 2025-05-09 Appendicies/AppendixB.tex
Added appendix B

\chapter*{Appendix B: Github Repositories}
\label{appendix:B}

This appendix provides links to the GitHub repositories associated with this thesis. These repositories contain the \LaTeX{} source files for the thesis and the code used for data handling and modeling.

\section*{Thesis \LaTeX{} Source Code}
The \LaTeX{} source code for this thesis can be found in the following GitHub repository:
\begin{itemize}
    \item \url{https://github.com/BrynjarGeir/lokaverkefni_vel_tex}
\end{itemize}

\section*{Data Handling and Modeling Code}
The code used for data handling and creating models, written in Python (in python files or notebooks), is available in the following GitHub repository:
\begin{itemize}
    \item \url{https://github.com/BrynjarGeir/lokaverkefni_vel}
\end{itemize}